---
title: "linear"
output: html_document
---

```{r setup, include=FALSE}
library(lmtest)
library(nortest)
library(dplyr)
library(e1071)
library(ISLR)
library(leaps)
library(VIF)
library(HH)
library(clusterGeneration)
library(MASS)
library(fmsb)
library(ggplot2)
options(stringsAsFactors = FALSE)
library(dplyr)
library(readr)
library(lubridate)
library(purrr)
library(ggplot2)
library(stringr)
library(data.table)
library(randomForest)
library(glmnet)
library(pls)
library(class)
library(gam)
library(akima)
library(nnet)
library(neuralnet)
library(NeuralNetTools)
library(ROCR)
library(leaps)
library(stats)
library(MASS)
library(qpcR)
library(MPV)
library(rgl)
library(MuMIn)
library(leaps)
library(bestglm)
library(qpcR)
library(rgl)
library(MuMIn)
library(bestglm)
library(forwards)
library(subselect)
library(FWDselect)
```

Data Cleaning
missing values
```{r}
loan<-read.csv("/Users/anmoe/Desktop/loan.csv", header=T)
dim(loan)
loan=na.omit(loan)
dim(loan)
```
emp_length as.numeric
```{r}
loan<-subset(loan, loan$emp_length!="n/a")
loan[which(loan$emp_length=='< 1 year'),]$emp_length<-0
loan[which(loan$emp_length=='1 year'),]$emp_length<-1
loan[which(loan$emp_length=='2 years'),]$emp_length<-2
loan[which(loan$emp_length=='3 years'),]$emp_length<-3
loan[which(loan$emp_length=='4 years'),]$emp_length<-4
loan[which(loan$emp_length=='5 years'),]$emp_length<-5
loan[which(loan$emp_length=='6 years'),]$emp_length<-6
loan[which(loan$emp_length=='7 years'),]$emp_length<-7
loan[which(loan$emp_length=='8 years'),]$emp_length<-8
loan[which(loan$emp_length=='9 years'),]$emp_length<-9
loan[which(loan$emp_length=='10+ years'),]$emp_length<-10
table(loan$emp_length)
loan$emp_length<-as.numeric(loan$emp_length)
```
fully paid:0     charged off:1
```{r}
a<-which(loan$loan_status=="Does not meet the credit policy. Status:Charged Off")
loan[a,]$loan_status<-c("Charged Off")
b<-which(loan$loan_status=="Does not meet the credit policy. Status:Fully Paid")
loan[b,]$loan_status<-c("Fully Paid")
levels(loan$loan_status) <- c(0,1)
loan$loan_status<-ifelse(loan$loan_status=="Fully Paid", 0, 1)
table(loan$loan_status)
loan$loan_status<-factor(loan$loan_status)
```
purpose: small_business:1   not small_business:0
```{r}
loan[which(loan$purpose!='small_business'),]$purpose<-0
loan[which(loan$purpose=='small_business'),]$purpose<-1
loan$purpose<-factor(loan$purpose)
```

```{r}
loan$grade<-factor(loan$grade)
loan$sub_grade<-factor(loan$sub_grade)
loan$home_ownership<-factor(loan$home_ownership)
loan$verification_status<-factor(loan$verification_status)
loan$purpose<-factor(loan$purpose)
loan$addr_state<-factor(loan$addr_state)
loan$debt_settlement_flag<-factor(loan$debt_settlement_flag)
#recoveries+total_rec_prncp+total_rec_int+total_rec_late_fee=total_pymnt
loan$total_rec_prncp<-NULL
loan$total_rec_int<-NULL
loan$total_rec_late_fee<-NULL
loan$recoveries<-NULL
#unknown until due day
loan$total_pymnt_inv<-NULL
#too much
loan$sub_grade<-NULL
```
loan: split badloan data into train data and test data
```{r}
set.seed(123)
loan.train = sample_frac(loan, 0.9)
sid<-as.numeric(rownames(loan.train))
loan.test<-loan[-sid,]
#badloan
dim(loan.train)
table(loan.train$loan_status)
dim(loan.test)
table(loan.test$loan_status)
```
badloan
```{r}
a<-which(loan.train$loan_status=="1")
badloan.train<-loan.train[a,]
b<-which(loan.test$loan_status=="1")
badloan.test<-loan.test[b,]
#write.csv(badloan,filee="/Users/anmoe/Desktop/badloan.csv")
badloan.train$loan_status<-NULL
dim(badloan.train)
```
regression reponse: total_pymnt
vif
```{r}
badloan.lm<-lm(total_pymnt~.,data=badloan.train)
vif<-vif(badloan.lm)
which.max(vif)
```

badloan regression: remove 4 predictors based on vif values (multicollinearity)
```{r}
badloan.train$funded_amnt<-NULL
badloan.train$addr_state<-NULL
badloan.train$loan_amnt<-NULL
badloan.train$grade<-NULL
#gradeE vif: 11.789794
```
regression: Forward Stepwise Selection: Choose 10 predictors
```{r}
badloan.fwd=regsubsets(badloan.train$total_pymnt~.,data=badloan.train,nvmax=51,method="forward")
reg.summary=summary(badloan.fwd)
par(mfrow=c(2,2))
plot(reg.summary$rss,xlab="Number of Variables",ylab="RSS",type="l")
plot(reg.summary$adjr2,xlab="Number of Variables",ylab="Adjusted RSq",type="l")
which.max(reg.summary$adjr2)
points(19,reg.summary$adjr2[19], col="red",cex=2,pch=20)
plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp",type='l')
which.min(reg.summary$cp)
points(17,reg.summary$cp[17],col="red",cex=2,pch=20)
plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(reg.summary$bic)
points(8,reg.summary$bic[8],col="red",cex=2,pch=20)
names(coef(badloan.fwd,8))
#[1] "(Intercept)"             "funded_amnt_inv"        
#[3] "term"                    "installment"            
#[5] "annual_inc"              "purposes1"  
#[7] "inq_last_6mths"          "collection_recovery_fee"
#[9] "debt_settlement_flagY"  
```
regression: selected data.trainn
```{r}
baddata<-cbind(badloan.train$total_pymnt,badloan.train$funded_amnt_inv,badloan.train$term,badloan.train$installment,badloan.train$annual_inc,badloan.train$purpose,badloan.train$inq_last_6mths,badloan.train$collection_recovery_fee,badloan.train$debt_settlement_flag)
colnames(baddata)<-c("y","funded_amnt_inv","term","installment","annual_inc","purpose","inq_last_6mths","collection_recovery_fee","debt_settlement_flag")  
baddata<-data.frame(baddata)

baddata.test<-cbind(badloan.test$total_pymnt,badloan.test$funded_amnt_inv,badloan.test$term,badloan.test$installment,badloan.test$annual_inc,badloan.test$purpose,badloan.test$inq_last_6mths,badloan.test$collection_recovery_fee,badloan.test$debt_settlement_flag)
colnames(baddata.test)<-c("y","funded_amnt_inv","term","installment","annual_inc","purpose","inq_last_6mths","collection_recovery_fee","debt_settlement_flag")  
baddata.test<-data.frame(baddata.test)
```

```{r}
#cor(baddata)
m0.lm<-lm(y~.,data=baddata)
summary(m0.lm)
lillie.test(m0.lm$residuals)
bptest(m0.lm, studentize = FALSE)
```
linear regression model for badloan.trainn
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
summary(m1.lm)$adj.r.squared

plot(m1.lm)
#0.5 normality  p-value: 0.03931  Adjusted R-squared:  0.5338
lillie.test(m1.lm$residuals)
bptest(m1.lm, studentize = FALSE)
#p-value: <2.2e-16
sum((m1.lm$fitted.values-sqrt(baddata$y))^2)/5543
```

weighted least regression w1<-1/y
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
w1<-1/m1.lm$fit
m2.lm<-lm(y^(0.5)~.,data=baddata, weights =w1)
summary(m2.lm) #Adjusted R-squared:  0.5365
anova(m2.lm) #MSE=8
plot(m2.lm)
lillie.test(m2.lm$residuals) #p-value = 6.14e-05
bptest(m2.lm, studentize = FALSE) #p-value < 2.2e-16
#curvlinear  quadr   log 
sum((m2.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m2.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)

```

weighted least regression w1<-1/res^2
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
w2<-1/(m1.lm$residuals^2)
m3.lm<-lm(y^(0.5)~.,data=baddata, weights =w2)
summary(m3.lm) #Adjusted R-squared:  0.9992
anova(m3.lm) #MSE=1
plot(m3.lm)
qqnorm(m3.lm$residuals)
lillie.test(m3.lm$residuals) #p-value = 0.03527
bptest(m3.lm, studentize = FALSE) #p-value < 2.2e-16
plot(baddata$y^(0.5))
#not include binary, see residuals 
#send email R squared 
#1/var   1/residual^2
sum((m3.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m3.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```

weighted least regression w1<-1/res^2   delete binary 
```{r}
m1.lm<-lm(y^(0.5)~baddata$funded_amnt_inv+baddata$installment+baddata$annual_inc+baddata$inq_last_6mths+baddata$collection_recovery_fee,data=baddata)
w2<-1/(m1.lm$residuals^2)
m3.lm<-lm(y^(0.5)~.,data=baddata, weights =w2)
summary(m3.lm) #Adjusted R-squared:  1
anova(m3.lm) #MSE=1
hist(m3.lm$residuals)
plot(m3.lm)
lillie.test(m3.lm$residuals) #p-value < 2.2e-16
bptest(m3.lm, studentize = FALSE) #p-value < 2.2e-16
plot(baddata$y^(0.5))
#not include binary, see residuals 
#send email R squared 
#1/var   1/residual^2

```



weighted least regression w1<-1/...
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
w3<-1/fitted(lm(abs(residuals(m1.lm)) ~ fitted(m1.lm)))^2
m4.lm<-lm(y^(0.5)~.,data=baddata, weights =w3)
summary(m4.lm) #Adjusted R-squared:  0.5289
anova(m4.lm) #MSE=1.5
plot(m4.lm)
lillie.test(m4.lm$residuals) #p-value = 7.692e-07
bptest(m4.lm, studentize = FALSE) #p-value < 2.2e-16
sum((m4.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m4.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```

weighted least regression w1<-1/abs(res)
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
w4<-1/abs(m1.lm$residuals)
m5.lm<-lm(y^(0.5)~.,data=baddata, weights =w4)
summary(m5.lm) #Adjusted R-squared:  0.8894
anova(m5.lm) #MSE=20
plot(m5.lm)
lillie.test(m5.lm$residuals) #p-value = 0.02514
bptest(m5.lm, studentize = FALSE) #p-value < 2.2e-16
sum((m5.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m5.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```
weighted least regression w1<-1/abs(res)
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
#externally studentized residuals
esr=studres(m2.lm)
w5<-1/abs(esr) #r^2:0.6947  normal p-val:0.011 non-constant p-val< 2.2e-16
m6.lm<-lm(y^(0.5)~.,data=baddata, weights =w5)
summary(m6.lm) #Adjusted R-squared:  0.9294
anova(m6.lm) #MSE=492
plot(m6.lm)
lillie.test(m6.lm$residuals) #p-value = 7.692e-07
bptest(m6.lm, studentize = FALSE) #p-value < 2.2e-16
sum((m6.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m6.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```
weighted least regression w1<-1/abs(weighted raw res)
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
#weighted raw residuals 
wres=weighted.residuals(m2.lm)
w6<-1/abs(wres) 
#weighted raw residuals 
m7.lm<-lm(y^(0.5)~.,data=baddata, weights =w6)
summary(m7.lm) #Adjusted R-squared:  0.9294
anova(m7.lm) #MSE=176
plot(m7.lm)
lillie.test(m7.lm$residuals) #p-value = 5.782e-05
bptest(m7.lm, studentize = FALSE) #p-value < 2.2e-16
#MSE
sum((m7.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m7.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```
robust regression psi=psi.bisquare
```{r}
m8.lm<-rlm(y^(0.5)~.,data=baddata, psi=psi.bisquare)
summary(m8.lm) #Residual standard error: 24.37 
#MSE=630.8849
anova(m8.lm)
n=5552
plot(baddata$y^(0.5)~fitted(m3.lm))
abline(0,1, col='red')
lillie.test(m8.lm$residuals) #p-value = 0.0003873
bptest(m8.lm, studentize = FALSE) #p-value < 2.2e-16
sum((m8.lm$fitted.values-sqrt(baddata$y))^2)/5543
pred=predict(m8.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)

```
robust regression psi=psi.hampel
```{r}
m9.lm<-rlm(y^(0.5)~.,data=baddata, psi=psi.hampel)
summary(m9.lm) #Residual standard error: 24.37 
#MSE=628.3365
anova(m9.lm)
n=5552
plot(baddata$y^(0.5)~fitted(m3.lm))
abline(0,1, col='red')
lillie.test(m9.lm$residuals) #p-value = 0.008273
bptest(m9.lm, studentize = FALSE) #p-value < 2.2e-16
pred=predict(m9.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```
robust regression psi=psi.huber
```{r}
m10.lm<-rlm(y^(0.5)~.,data=baddata, psi=psi.huber)
summary(m10.lm) #Residual standard error: 24.37 
#MSE=629.5456
anova(m10.lm)
n=5552
plot(baddata$y^(0.5)~fitted(m10.lm))
abline(0,1, col='red')
lillie.test(m10.lm$residuals) #p-value = 0.001817
bptest(m10.lm, studentize = FALSE) #p-value < 2.2e-16
pred=predict(m10.lm,newdata=baddata.test,se=F)
sum((pred-sqrt(baddata.test$y))^2)/(767-8)
```
delete influtial cases   NO NEED
```{r}
m1.lm<-lm(y^(0.5)~.,data=baddata)
#5552 observations in baddata
im<-influence.measures(m1.lm)
im$is.inf[1,]==F
baddata_influential_obs <- which(apply(im$is.inf, 1, any)) 
baddata_delete_influential_obs <- baddata[-baddata_influential_obs,]
#5011 observations in baddata
m11.lm<-lm(y^(0.5)~.,data=baddata_delete_influential_obs)
summary(m11.lm)
plot(m11.lm)
#0.5 normality  p-value: 0.0002574  Adjusted R-squared:  0.5759
lillie.test(m11.lm$residuals)
bptest(m11.lm, studentize = FALSE)
summary(m1.lm)
lillie.test(m1.lm$residuals)
bptest(m1.lm, studentize = FALSE)
```
delete influtial cases   NO NEED
```{r}
m1.lm<-lm(y^(0.5)~installment+collection_recovery_fee+funded_amnt_inv:term+funded_amnt_inv:installment+funded_amnt_inv:purpose+funded_amnt_inv:inq_last_6mths+funded_amnt_inv:debt_settlement_flag+term:installment+term:annual_inc+term:collection_recovery_fee+installment:debt_settlement_flag+collection_recovery_fee:debt_settlement_flag,data=baddata)
#5552 observations in baddata
im<-influence.measures(m1.lm)
im$is.inf[1,]==F
baddata_influential_obs <- which(apply(im$is.inf, 1, any)) 
baddata_delete_influential_obs <- baddata[-baddata_influential_obs,]
#4995 observations in baddata
m11.lm<-lm(y^(0.5)~installment+collection_recovery_fee+funded_amnt_inv:term+funded_amnt_inv:installment+funded_amnt_inv:purpose+funded_amnt_inv:inq_last_6mths+funded_amnt_inv:debt_settlement_flag+term:installment+term:annual_inc+term:collection_recovery_fee+installment:debt_settlement_flag+collection_recovery_fee:debt_settlement_flag,data=baddata_delete_influential_obs)
summary(m11.lm)
plot(m11.lm)
#0.5 normality  p-value: 0.0002574  Adjusted R-squared:  0.5759
lillie.test(m11.lm$residuals)
bptest(m11.lm, studentize = FALSE)
summary(m1.lm)
lillie.test(m1.lm$residuals)
bptest(m1.lm, studentize = FALSE)
qchisq(0.95,12)
```
(interaction, higher order) subset selection for baddata
```{r}
baddata.fwd=regsubsets(y^(0.5)~funded_amnt_inv+term+installment+annual_inc+purpose+inq_last_6mths+collection_recovery_fee+debt_settlement_flag
+funded_amnt_inv^2+term^2+installment^2+annual_inc^2+inq_last_6mths^2+collection_recovery_fee^2 +funded_amnt_inv*term+funded_amnt_inv*installment+funded_amnt_inv*annual_inc+funded_amnt_inv*purpose+funded_amnt_inv*inq_last_6mths+funded_amnt_inv*collection_recovery_fee+funded_amnt_inv*debt_settlement_flag+term*installment  +term*annual_inc+term*purpose +term*inq_last_6mths+term*collection_recovery_fee  +term*debt_settlement_flag+installment*annual_inc+installment*purpose  +installment*inq_last_6mths+installment*collection_recovery_fee +installment*debt_settlement_flag+annual_inc*purpose+annual_inc*inq_last_6mths +annual_inc*collection_recovery_fee+annual_inc*debt_settlement_flag+purpose*inq_last_6mths+purpose*collection_recovery_fee +purpose*debt_settlement_flag+inq_last_6mths*collection_recovery_fee+  inq_last_6mths*debt_settlement_flag+collection_recovery_fee* debt_settlement_flag,data=baddata,nvmax=51,method="forward")
reg.summary=summary(baddata.fwd)
plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
points(12,reg.summary$bic[12],col="red",cex=2,pch=20)
which.min(reg.summary$bic)
names(coef(baddata.fwd,12))
#[1] "(Intercept)"                                 
# [2] "installment"                                 
# [3] "collection_recovery_fee"                     
# [4] "funded_amnt_inv:term"                        
# [5] "funded_amnt_inv:installment"                 
# [6] "funded_amnt_inv:purpose"                     
# [7] "funded_amnt_inv:inq_last_6mths"              
# [8] "funded_amnt_inv:debt_settlement_flag"        
# [9] "term:installment"                            
#[10] "term:annual_inc"                             
#[11] "term:collection_recovery_fee"                
#[12] "installment:debt_settlement_flag"            
#[13] "collection_recovery_fee:debt_settlement_flag"
```

linear regression model for badloan.train sqrt(y)
```{r}
m2.lm<-lm(y^(0.5)~installment+collection_recovery_fee+funded_amnt_inv:term+funded_amnt_inv:installment+funded_amnt_inv:purpose+funded_amnt_inv:inq_last_6mths+funded_amnt_inv:debt_settlement_flag+term:installment+term:annual_inc+term:collection_recovery_fee+installment:debt_settlement_flag+collection_recovery_fee:debt_settlement_flag,data=baddata)
summary(m2.lm)
anova(m2.lm)
#Adjusted R-squared:  0.5574
lillie.test(m2.lm$residuals)
#normality p-value: 0.001138
```
```{r}
m2.lm<-lm(y^(0.5)~installment+collection_recovery_fee+funded_amnt_inv:term,data=baddata)
anova(m2.lm)
qf(0.95,1,5548)
```
delect insignificant term 
```{r}
m3.lm<-lm(y^(0.5)~installment+collection_recovery_fee+funded_amnt_inv:term+funded_amnt_inv:installment+funded_amnt_inv:purpose+funded_amnt_inv:inq_last_6mths+term:installment+term:annual_inc+term:collection_recovery_fee+installment:debt_settlement_flag+collection_recovery_fee:debt_settlement_flag,data=baddata)
summary(m3.lm)
#Adjusted R-squared:  0.5574
lillie.test(m3.lm$residuals)
#normality p-value: 0.0006279
```



```{r}

```








```{r}

```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

